### 2023-04-06 main.tex

\documentclass{article}

\usepackage[english]{babel}
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

\usepackage{authblk}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{makecell}

\title{Towards a Pandemic-Scale Ancestral Recombination Graph of SARS-CoV-2}
\author[1]{Shing H. Zhan}
\author[2,5]{Katherine Eaton}
\author[3,5]{Anastasia Ignatieva}
\author[1,5]{Benjamin Jeffery}
\author[1,5]{Duncan S. Palmer}
\author[1,5]{Yan Wong}
\author[2]{Carmen L. Murall}
\author[4]{Sarah P. Otto}
\author[1,6]{Jerome Kelleher}
\affil[1]{Big Data Institute, Li Ka Shing Centre for Health Information and Discovery, University of Oxford, Oxford, United Kingdom}
\affil[2]{National Microbiology Laboratory, Public Health Agency of Canada, Canada}
\affil[3]{Department of Statistics, University of Oxford, Oxford, United Kingdom}
\affil[4]{Department of Zoology and Biodiversity Research Centre, University of British Columbia, Vancouver, British Columbia, Canada}
\affil[5]{Preliminary alphabetic ordering}
\affil[6]{Correspondence. E-mail: jerome.kelleher@bdi.ox.ac.uk}

\begin{document}
\maketitle

\begin{abstract}
XX
\end{abstract}

\section{Introduction}

Recombination via template switching is a common feature in the evolution of coronaviruses (CoV) \cite{Graham2010-xe,De_Klerk2022-tt}), including SARS-CoV-2 \cite{VanInsberghe2021-eu,Jackson2021-ik,Ignatieva2021-rg}. By bringing together mutations carried by different lineages, recombination plays an important role in generating genomic diversity, with recombinant lineages associated with CoV adaptation to new host species and with the production of more immune evasive variants \cite{Graham2010-xe,De_Klerk2022-tt}.

Recombination was challenging to detect in SARS-CoV-2 early in the pandemic due to gradual accumulation of genetic variation. In late 2020, the appearance and spread of variants of concern (VoC) harbouring multiple characteristic mutations increased the detectability of recombination. By searching for genomes containing segments from different VoC, multiple recombinant lineages have been detected, for example, between Alpha and pre-VoC \cite{Jackson2021-ik}, Alpha and Epsilon \cite{Wertheim2022-hj}, and Alpha and Delta \cite{Sekizuka2022-xz}). More recently, the high prevalence of Omicron, with multiple co-circulating deeply divergent lineages (BA.1 to BA.5), has accelerated the rate of coinfection and the potential for recombination \cite{Bal2022-hq}. Recombinant lineages that have sufficient evidence of onward transmission (more than five descendants) and monophyly (> 70\% bootstrap support) are eligible for a Pango lineage designation \cite{Rambaut2020-dw}, starting with the letter X (“Pango Lineage Nomenclature: Provisional Rules for Naming Recombinant Lineages” 2021).

Most genomic surveys for SARS-CoV-2 recombinants search for mosaic genomes that combine specific subsets of mutations from different lineages (e.g., \cite{VanInsberghe2021-eu,Jackson2021-ik,Wertheim2022-hj,Sekizuka2022-xz}). This approach only identifies those inter-lineage recombination events resulting in the patterns investigated. Others have inferred recombinant genealogies to disentangle recombination from recurrent mutation, without having to pre-specify characteristic mutations for each lineage \cite{Ignatieva2021-rg}. More recently, Turakhia et al. \cite{Turakhia2022-it} has presented a phylogeny-based approach ("RIPPLES") to identify putative recombinants among over ten million SARS-CoV-2 genomes, without pre-specifying motifs. To find candidate recombinants, RIPPLES first builds a phylogeny without assuming recombination and then scans for long branches (i.e., carrying many mutations) and then determines if these candidates would be better explained by recombination. Specifically, RIPPLES exhaustively breaks each candidate sequence into segments and attempts to find more parsimonious placements for each segment on the phylogeny. If such placements are found, the sequence is identified as a putative recombinant. Although it enables rapid searches for genomic evidence of recombinants, RIPPLES relies on a SARS-CoV-2 phylogeny that accounts for only mutations, treats recombinants \textit{post hoc}, and is an incomplete representation of the reticulate evolutionary history of SARS-CoV-2. As noted by the authors, a \textit{post hoc} treatment of recombination is possible when recombinant lineages are rare and leave few descendants, but this is becoming increasingly untenable with the proliferation of recombinant lineages (e.g., more than half of the sequences sampled in February, 2023 are from XBB and its descendants; \cite{Chen2022-pz}.

Inferring evolutionary trees while accounting for both mutation and recombination is a long-standing challenge. This challenge is particularly acute for SARS-CoV-2, given the large amount of data that has been shared, with over 15 million sequences in the GISAID database \cite{Shu2017-hp}. It is well known that recombination distorts phylogenies \cite{Schierup2000-fg} and affects the results of downstream analysis (e.g., inference of selection; \cite{Anisimova2003-vr}). However, standard phylogenetic methods do not account for recombination (e.g., \cite{Ronquist2012-zw,Minh2020-lr,Guindon2003-zd}). Alternatively, genetic ancestries of recombining lineages can be represented by ancestral recombination graphs (ARG; \cite{Griffiths1981-lw,Hudson1983-if}, which  capture network-like phylogenetic relationships \cite{Gusfield2014-qw}.

Approaches to infer ARGs now exist that can scale to tens of thousands of human genomes and beyond \cite{Speidel2019-yh,Kelleher2019-ba,Zhang2021-lf,Schaefer2021-yg}. Recently, Wohns et al. \cite{Wohns2022-th} reconstructed a unified whole-genome genealogy of thousands of modern and ancient human genomes, based on over 90 million variable sites. To achieve this, Wohns et al. \cite{Wohns2022-th} utilised several computational advances in ARG inference: (1) the "succinct tree sequence" data structure that can compactly store large ARGs  \cite{Kelleher2018-xc}; (2) a highly efficient Hidden Markov Model (HMM) implementation of the Li and Stephens (LS) model (Li and Stephens 2003) to infer copying paths between genomes \cite{Kelleher2019-ba}; and (3) efficient, open-source software tools \cite{Baumdicker2022-ep,Kelleher2018-xc,Kelleher2019-ba,Wohns2022-th}.

Here, we present a preliminary version of a novel scalable methodology (“sc2ts”, pronounced “scoots”) that builds on the advances above and additionally takes advantage of known collection dates to infer ARGs from millions of full-length SARS-CoV-2 genomes. Using publicly available SARS-CoV-2 genome sequences, we demonstrate that sc2ts can be used to simultaneously (1) detect putative SARS-CoV-2 recombinants; (2) infer the phylogenetic relationships of the major SARS-CoV-2 lineages; (3) reconstruct the mutational spectrum of SARS-CoV-2; and (4) provide a compact record of evolutionary information about SARS-CoV-2 that can be efficiently accessed using "tskit" (tree sequence toolkit) software library \cite{Kelleher2018-xc}. Sc2ts is a promising method to automate detection of SARS-CoV-2 recombinants and to jointly model recombinational and mutational processes shaping the genomic diversity of SARS-CoV-2 at large scale.

\section{Methods}

\subsection{Data preprocessing}

\subsubsection{Sequence-level preprocessing}

The findings of this study are based on sequences and metadata available on GISAID (\href{https://gisaid.org/}) up to Aug 22, 2022 and accessible at doi.org/10.55876/gis8.230329cd (Supplementary Table 1). Sequences were removed if they (1) had ambiguous collection dates; (2) were collected before January 01, 2020 or after June 30, 2022; or (3) were isolated from a non-human host. Sequences were aligned to the Wuhan-Hu-1/2019 reference sequence (GenBank: MN908947.3) using Nextclade v2.3.0 \cite{Aksamentov2021-hj} (dataset tag 2022-07-26T12:00:00Z). To reduce the impact of poor-quality sequences on ARG inference, sequences were excluded if they had a “bad” quality control status in any of the four Nextclade columns ("qc.missingData.status”, “qc.mixedSites.status”, “qc.frameShifts.status”, and “qc.stopCodons.status”). These excluded genomes are enriched for low-coverage, contaminated, or misassembled sequences.

\subsubsection{Site-level preprocessing}

Ambiguous nucleotide letters (i.e., not A, C, G, T, or a gap) in the pairwise genome alignments were encoded as N. Problematic bases in the alignments, which had two or more ambiguous letters or gaps within a distance of seven bases, were masked as missing data following the protocol in \href{https://github.com/roblanf/sarscov2phylo/}. Masked bases were ignored when attaching a sample genome to the ARG, but they were subsequently imputed. Further, 481 problematic sites flagged as prone to sequencing errors or highly homoplasic were excluded entirely (\href{https://virological.org/t/issues-with-sars-cov-2-sequencing-data/473}; \href{https://raw.githubusercontent.com/W-L/ProblematicSites_SARS-CoV2/master/problematic_sites_sarsCov2.vcf}).

\subsubsection{Exclusion of “time travellers”}

Including samples with incorrectly reported collection dates (e.g., an Alpha sample purportedly collected in 2020 from the United States, before Alpha appeared in the United Kingdom, USA/MN-Mayo-1563/2020) was found to produce artefacts that distorted the topology of ARGs inferred early on in the study. Hence, to exclude such potential “time travellers”, we employed two filters. The first was to remove any sequence with a collection date that pre-dated the time to the most recent common ancestor (tMRCA) of its corresponding clade. The tMRCA for each clade was obtained from a Nextstrain GISAID reference tree (downloaded on August 22, 2022) where we used the lower bound of the 95\% confidence interval of each clade as the minimum date cut-off. The second filter was a simple threshold on the time delay between the collection date and submission date (maximum submission delay, or MSD). Indeed, we found that ARGs inferred from the MSD-filtered genome set had noticeably fewer branches with a high number of reversions. After some preliminary analysis, a MSD threshold value of 30 days was used when building the ARGs described herein. Future work could estimate collection dates by adding these potential time travellers back on the tree, allowing an automated assessment of collection date discrepancies.

\subsection{Inference}

\subsubsection{Overview of sc2ts}

We introduce an ARG inference method (schematic in Figure 1) that uses (1) an efficient HMM implementation of the Li and Stephens (LS) model \cite{Li2003-ib} to copy SARS-CoV-2 genomes (see Sections 2.2 and 2.3); (2) the reported collection dates of the SARS-CoV-2 samples to determine the order of genome copying; and (3) several parsimony-based heuristics to locally refine the topology. In brief, this combined approach uses the likelihood model of LS to attach sample genomes to existing nodes and to detect recombinants, but because many of the tip samples attached daily harbour few, if any, mutational changes, we use parsimony to resolve the relationships near the tip samples.

Our HMM implementation, sc2ts, builds a reticulated graph for SARS-CoV-2 by taking advantage of the reported collection dates to incrementally grow the ARG. The inference process occurs forward in time, attaching new genomes aggregated by their collection date (daily batches), from the earliest to the last collection date. Prior to attaching sample genomes, the ARG is initialised such that the Wuhan-Hu-1/2019 reference genome is the root node. In the absence of recombination, the ARG is a single tree with no reticulation (Figure 1A); however, in the presence of recombination, a reticulate ARG is generated, with variation in tree topology across the genome (Figure 1B). Unlike standard phylogenetic trees in which internal nodes represent inferred ancestors, internal nodes in the ARGs inferred using sc2ts represent ancestors which can be samples or inferred ancestors.

Each daily batch of genomes is first attached to the growing ARG without any phylogenetic resolution. In cases where the newly added sequence is inferred to be a recombinant, a recombination node is inserted; the “date” of this node is arbitrarily set to the average of the collection dates of the genome and the most recent of its parent nodes. Otherwise, the genomes are attached to the most closely related ancestral node (an attachment node). This creates polytomies that are then locally resolved among the descendants of each attachment node using the Unweighted Pair Group Method with Arithmetic Mean (UPGMA) algorithm \cite{Michener1957-tr}, with average linkage (Figure 1C). This hierarchical clustering algorithm produces a dendrogram relating a set of sequences, which is treated as an ultrametric binary tree here. Subsequently, mutations are placed onto the daily tree by maximum parsimony. An issue with the UPGMA algorithm is that it may create internal nodes supported by no informative site (i.e., having no mutation immediately ancestral to them). In an adjustment step, such unsupported internal nodes are removed to “un-resolve” the daily tree; the remaining internal nodes, which are supported by at least one shared mutation, are retained. The “dates” of the non-sample internal nodes are set to the same date as the samples in the daily batch. The resulting daily tree is grafted onto the attachment node in the global ARG.

During ARG inference, two types of non-sample nodes are inserted to improve parsimonious placement of mutations: mutation-coalescing nodes and pushed-reversion nodes. We call these operations tree adjustments. A mutation-collapsing node is inserted when a shared mutation between sibling nodes is found (Figure 1D). An attempt to create mutation-collapsing nodes is made after grafting a new daily tree onto the global ARG. If shared mutations exist between two or more sibling nodes, then a parsimonious explanation is to create a new node that is the parent of those siblings, on which these mutations occurred. A pushed-reversion node is inserted when an immediate reversion occurs (Figure 1E), i.e. a mutation that reverts another mutation on its parent node. Instead of retaining both the mutation and immediate reversion, only the mutation is kept after a pushed-reversion node is inserted, thereby reducing the number of mutations on the branches. Note that these non-sample nodes are used as reconstructed genomes during HMM copying.

Furthermore, during inference, the ARG is automatically encoded, stored, and updated as a tree sequence using tskit \cite{Kelleher2018-xc}.

\subsubsection{Brief description of the Li and Stephens model}

The LS model is a convenient probabilistic approximation to the coalescent with recombination \cite{Hudson1983-if}, itself a generative model of a genealogical process with recombination \cite{Li2003-ib}. Importantly, it can be implemented as an HMM in which a focal genome is a sequence of nucleotides (observed states) that are probabilistically emitted as an imperfect mosaic of a set of reference genomes (hidden states). As such, an efficient HMM machinery can be employed to make statements about parameters of interest and highlight potential sources of error in genotyping and sequencing data. The LS model is used in a wide variety of applications in genomics. For example, it is the engine that drives statistical genotype phasing and imputation \cite{Delaneau2019-wl,Browning2021-cg,Browning2018-nk,Rubinacci2020-pa}, and it can be used to estimate mutation rates (e.g., REF), recombination rates (e.g., \cite{Hinch2011-tz}), and selection within and across hosts in viral sequence data (e.g., \cite{Palmer2019-wa}).

The generative process of the LS model is summarised in Figure 2. Here, a transition matrix, Q, governs the process of switching (recombining) between members of the reference panel (the hidden states). An emission matrix, \textit{E}, allows for differences between the focal sequence and the hidden state from which it is copied (due to mutation as well as genotyping or sequencing error). Then, we use the Viterbi algorithm \cite{Viterbi1967-ol} to find the most likely copying path through the sequences, given a set of reference sequences, \textit{Q}, and \textit{E}.

Because the HMM is memoryless, the potential parental lineages are independently assessed at each recombinantion breakpoint. This can result in the inference of a recombinant lineage descending immediately from three (or even more) parental lineages. We find that often two of the three parents are identical or similar in sequence in the segments contributed to the recombinant lineage. A more parsimonious explanation would be to switch back to the original parent sequence. In the future, a pragmatic solution would be to detect such cases during the ARG inference stage and to adjust the copying paths accordingly. Meanwhile, however, we urge caution when interpreting the parentage of lineages with multiple breakpoints.

\subsubsection{Parameterization of the HMM implementation of Li and Stephens model}

Recombination is inferred using an efficient HMM implementation of the LS model. Copying paths are estimated from SARS-CoV-2 genomes sampled over time, with younger genomes copying (i.e. inheriting segments and mutations) from older genomes. Here, we were able to apply the LS model on millions of SARS-CoV-2 genomes because of a highly efficient, exact HMM implementation of the LS model that is a core component of the tsinfer software. For this application, we have parameterised the HMM using a mismatch ratio (MMR), which tunes how strongly mutations (i.e., no switching of template sequences) are preferred over recombination (i.e., switching between sequences) during copying. Under a lower MMR value, the maximum-likelihood HMM copying path of a focal genome more likely spans at least two ancestral genomes (i.e,. the focal genome is a putative recombinant) than a single ancestral genome. Here, the MMR was set to three in all the HMM runs (see Discussion for future potential development of the HMM).

\subsection{Analysis}

\subsubsection{Imputation of Pango lineage for non-sample nodes}

Non-sample nodes inserted during ARG inference had no Pango lineage assignment from Nextclade. We imputed the Pango lineage of non-sample nodes using a two-step procedure (Figure 3) to help assess how consistently breakpoint intervals are estimated by the HMM runs (see below). If the Pango lineage of the parental node was available and no mutation was immediately ancestral to the non-sample node, then the parental node’s status was directly propagated to the non-sample node. Otherwise, the Pango lineage of the non-sample node was inferred by matching its full set of mutations against the Pango lineage-defining mutations (based on 90\% consensus of the sequences analysed) from the COVID-CG website \cite{Chen2021-zc}; \href{https://covidcg.org/}; accessed on November 04, 2022).

\subsubsection{Breakpoint intervals and HMM consistency}

When recombinant ancestry for a given genome is inferred by the LS HMM, a breakpoint is identified as the last possible position where the copying path must switch. Thus, the breakpoints used in the ARG represent the rightmost end of the interval, within which the recombination event could have occurred. To obtain the leftmost end of this interval, we rerun the HMM using mirrored coordinates (i.e., from the 3’ end to the 5’ end).

A recombinant sequence was deemed to be HMM-consistent if (1) the same number of parent nodes and (2) the same set of mutations immediately ancestral to it were obtained from both the unmirrored and mirrored HMM runs. In the ideal case, the copying paths of both the HMM runs switch between parent sequences at the same position(s); in less ideal cases, however, the copying paths switch at different positions, typically because of a lack of genetic variation between the positions. The switch positions in the copying paths of both the HMM runs were used to define the breakpoint interval(s) of an HMM-consistent recombinant.

\subsubsection{Code availability}

The source code implementing the sc2ts pipeline is available via GitHub \href{https://github.com/jeromekelleher/sc2ts/}. The sc2ts pipeline uses various functions and algorithms implemented in tskit and tsinfer \href{https://github.com/tskit-dev/}. The pipeline uses the SciPy implementation of the UPGMA algorithm to build daily trees \href{https://docs.scipy.org/doc/scipy/reference/generated/scipy.cluster.hierarchy.average.html}. The Jupyter notebooks used to produce the results described herein are also available on GitHub \href{https://github.com/jeromekelleher/sc2ts-paper/}.

\section{Results}

\subsection{Inferred ARGs}

To illustrate the use of sc2ts, we employed two different sampling strategies. First, we inferred a densely sampled but time-limited “wide” ARG by attaching all the sample genomes with an MSD of 30 days, one daily batch at a time. The inference procedure ran until the genomes sampled on June 30, 2021 were attached to the ARG, resulting in a total of 1.27 million samples. Second, we inferred a “long” ARG over a more extended timespan, by attaching at most 1,000 genomes, which were randomly subsampled from each daily batch with an MSD within 30 days. For this ARG, the inference procedure ran until the genomes sampled on June 30, 2022 were attached to the ARG. The long ARG covers an extra year of samples but is more sparsely sampled. June 30, 2022 is the collection date of the final batch of sequences which were preprocessed for the development of sc2ts. Once sc2ts is further developed and tested, we will attach samples collected after that date to infer a fully sampled pandemic-scale ARG.

Most of the nodes in the ARGs represent sample genomes (Table 1). Only a small fraction of the nodes represent recombination events (0.28\% and 0.27\% in the wide ARG and long ARG, respectively), over half of which have no descendant samples (besides the recombinant sequence that triggered insertion of the recombination node) (Table 1). There are over one million mutations on the ARGs, the majority of which occur on a terminal edge (Table 1). Notably, the massive amount of genomic variation data in the wide ARG (1.27 million genomes) and long ARG (657 thousand genomes) are compactly stored in ~57 Mb and ~36 Mb, respectively (less than ten bytes per sample). These ARGs can be loaded for analysis in a Jupyter notebook on a typical commercial laptop in merely a fraction of a second using the Python package tskit (Table 1).

\subsection{Computational resources}

The wide ARG (with over one million sample genomes) was built using 128 threads on a machine with 512 GB RAM (2x AMD EPYC 7502 @ 2.5GHz). Inference ran for 17 days to attach genomes sampled up to June 30, 2021. The long ARG (with ~700 thousand sample genomes) was built using 40 threads on a machine with 256 GB RAM (2x Intel(R) Xeon(R) CPU E5-2680 v2 @ 2.80GHz). Inference ran for 23 days to attach genomes sampled up to June 30, 2022. Most of the runtime was spent on running the HMM.

\subsection{Backbone phylogeny}

For validation purposes, we compare the genetic relationships in the sc2ts-inferred ARGs with phylogenetic relationships inferred by the NextStrain project \cite{Hadfield2018-ef}. The NextStrain trees consist of subsamples of SARS-CoV-2 genomes, which over the time period of our ARGs are almost entirely restricted to non-recombinant lineages. To allow for comparison, the ARGs were "simplified" \cite{Kelleher2018-xc} to retain only those tips of our genealogy that match the samples present in a specific NextStrain tree. These simplified "backbone" ARGs exhibit very little recombination (6 and 5 trees across the genome for the wide and long ARGs, respectively).

Figure 4 shows a co-phylogeny of the 88 samples shared between the wide ARG and a GISAID global "all-time" tree from NextStrain (downloaded on January 21, 2023). To reduce the ARG to a single tree, we show the very first tree in the wide ARG; however, the five other trees in this simplified ARG show near-identical patterns. Supplementary Figure 3 shows the same comparison for the long ARG. It is clear that the backbone topology of the sc2ts tree shows very close agreement with the NextStrain tree. The sample genomes cluster by their assigned Pango lineage status, and many variants and their descendents form identical monophyletic clades in both the trees (e.g., the Alpha and Delta clades).

There are, however, some notable differences between the trees. Firstly, the sc2ts tree is generally less well resolved (noticeably, February to March, 2020) (Figure 4). Secondly, there are non-identical sample partitions near the tips (e.g., in the Alpha clade, Figure 4). These might have arisen because the samples containing the mutations needed to resolve local topologies were collected later than the cut-off date for the sc2ts ARG, but were for the NextStrain tree. Finally, the branch lengths differ. For example, the branch leading to the ancestral node of the Delta variant samples is much shorter in the sc2ts tree than the NextStrain tree (Figure 4). It is possible that the length of some of the branches was poorly calibrated, probably caused by overly aggressive insertion of mutation-collapsing nodes and reversion-pushed nodes or suboptimal sample attachment. Side-by-side comparison of sc2ts trees with trees inferred using another method (e.g., UShER) should help to understand whether branch length differences are due to poor calibration or methodological differences.

Although here we focus on a single local tree at an arbitrarily chosen position, note that in the original “un-simplified” ARGs, there are 1,496 and 958 trees, respectively (Table 1). Hence, while for a relatively small sample the genealogy may be well described by a single tree, the full dataset shows substantial variation in phylogenetic relationships (or phylogenetic incongruence) across the SARS-CoV-2 genome, much of which may be attributable to recombination. Such complex reticulate relationships are appropriately captured by a genealogy in the form of an ARG.

\subsection{Putative recombination events}

Both the wide and long ARGs contain a large number of recombination nodes (4,123 and 2,078, respectively; Supplementary Tables 1 and 2), with each of these nodes inferred as a result of matching a specific sample (a recombinant) into the ARG. However, ~65\% of recombination nodes are ancestors of only a single descendant sample in each local tree; these could have arisen from sequencing of mixed samples (e.g., from co-infection or contamination), or were simply not transmitted onwards and so were excluded from subsequent analyses (leaving X and Y recombinants in the wide and long ARGs, respectively). In their analysis using RIPPLES, Turakhia et al. (2022) detected 589 recombinants, also excluding singletons.

For simplicity, we focus primarily on “non-complex” recombination nodes. To do this, we apply a filtering step that removes those with more than one breakpoint (and often more than two parent lineages), or where matching is not HMM-consistent or identifies parents with different Pango lineage assignments in the original (i.e., 5’ to 3’ end) and mirrored (i.e., 3’ to 5’ end) HMM runs. Filtering reduces both the number of recombination nodes and breakpoints from  A \& B to C in the wide ARG and from 726 & 806 to 459 in the long ARG (Supplementary Tables 1 and 2). Note that in the unfiltered case, a single recombination node can correspond to multiple points, one for each recombination breakpoint. The effect of this filtering step in the long ARG, which includes time periods of increased rates of coinfection and recombination compared to the wide ARG, can be seen in Figure 3A (unfiltered) and Figure 3B (filtered).

Importantly, the long ARG captured all except one of the Pango designated recombinant lineages (see the section below), but we also detected a large number of additional recombination events, with parents lineages that diverged as far back as two years to as recent as five or six days prior to the recombination event (Figure 3). This provides a degree of resolution of viral recombinants not available with other recombination detection methods. The majority of newly identified recombinants arise from parents that diverged 10 to 30 weeks previously. However, there is also a peak of recombinants that diverged ~95 weeks prior to the event: these correspond to recombinants whose parental common ancestor traces to early 2020, including both Delta-Delta and Delta-Omicron recombinants (Supplementary Figure 4).

The date of common ancestry for the two parents of recombinant lineages is concentrated in several banded rows in Figure 5. These are largely due to a small handful of most recent common ancestors (MRCAs), which are shared between a large number of recombination events, the top four of which are indicated by dotted lines in the figure. These MRCAs are near the root of large expansions of certain lineages (B.1: original variant;  B.1.617.2: Delta, BA.1: first Omicron wave; and BA.2: second Omicron wave), as evidenced by the number of samples with that Pango designation that trace back to these ancestral nodes (listed as percentages on the plot). The specific MRCA nodes to which many recombinants trace frequently represent very large polytomies in the tree (at the extreme, 2,872 immediate children). This is likely to indicate a rapid and under-sampled expansion of a clade in the SARS-CoV-2 genealogy.

Finally, it is clear from Supplementary Figure 4 that of the recombinant parents that can be classified into Alpha, Delta, and Omicron variants, the majority are within-Delta and within-Omicron recombinants, and that Omicron followed by Delta, are the variants associated with the most recombination. This may reflect either sampling intensity, the prevalence of cases (which increases the chance of coinfection and recombination), or possibly heterogeneity in recombination rates among lineages.

\subsection{Early recombinants reported by Jackson et al. (2021)}

By searching for samples combining genomic segments from Alpha (B.1.1.7) and from the parental lineage B.1.1 based on a list of 22 Alpha-defining mutations, Jackson et al. (Jackson et al. 2021) reported 16 recombinant sequences among genomes sampled in the United Kingdom up to March 7, 2021. The authors proposed that the recombinants likely arose from eight independent recombination events (categorised into groups A to D and four singletons).

In the wide ARG (sampled to mid-2021), all these sequences are descendants of recombination nodes, except MILK-103C712, which was removed during preprocessing (Table 2); additionally, all are HMM-consistent, except QEUH-1067DEF. Overall, we find excellent agreement between Jackson et al.’s motif-based method and sc2ts in terms of (1) the Pango lineage of the parents and (2) the breakpoint intervals (Table 2). In all cases, the two methods proposed identical or very closely related parent lineages. The general mosaic genome structure is concordant in all the cases, even CAMC-CB7AB3 where there are two switches in the HMM copying paths (indicative of two breakpoints). The breakpoint intervals estimated using the two methods overlap substantially in all the cases (Table 2). Within each of the groups A to D, the constituent sequences have identical breakpoint intervals, because they trace back to a single recombination node in the wide ARG. This result is consistent with the hypothesis that the sequences of the groups A to D emerged from four separate recombination events (Jackson et al. 2021).

\subsection{Pango recombinants}

The wide ARG only contains samples from three Pango-designated recombinant lineages (XA, XB, and XC), due to the fact that it only includes samples up to mid-2021. By contrast, the long ARG (subsampled to mid-2022) contains samples from 33 Pango-designated recombinant lineages (XA to XAK; Supplementary Table 3). For Pango recombinant analysis, we therefore focus on the long ARG.

In the wide ARG, every sequence of the three Pango recombinants trace back to some recombination node. XA, XB, and XC.

In the unfiltered long ARG, the sequences of 28 Pango recombinants trace back to some recombinant node(s); however, the sequences of the five Pango recombinants XP, XU, XAA, XAG, and XAK do not. A likely explanation for the XP recombinant sequences not having a recombinant origin in the long ARG is that its characteristic multibase deletion (https://github.com/cov-lineages/pango-designation/issues/481) was masked during preprocessing and therefore was not incorporated into the HMM.
Upon tracing the origin of XAG through the ARG, XAG might have been caused by incorrect imputation of its parent nodes(?). XAK likely does not appear to have a recombinant origin because of random subsampling, but it may with a more complete tree.

The sequences of 14 Pango recombinants trace back to a single recombination node. The sequences of the remaining 18 Pango recombinants trace back to multiple recombination nodes, however. This may be a result of undersampling in the long ARG, which was built by randomly subsampling at most 1,000 sequences from each daily batch. Having extra recombination nodes in the long ARG appeared to be a more parsimonious explanation for the 19 Pango recombinants than having a single recombination node with many immediately ancestral mutations (including recurrent mutations and reversions), despite that the recombinants may indeed trace to a single origin.

The mosaic genome structure of all the 32 Pango recombinants is concordant with the structure determined by the Pango Lineage Designation Committee (Supplementary Table 3). For example, the genome of XD (also known as “Deltacron”) is inferred to consist of B.1.617.2 (Delta) on the 5’ end, BA.1.17 (Omicron) in the middle, and B.1.617.2 (Delta) on the 3’ end.

\subsection{Novel recombinants}

Next, we examine recombination nodes in the filtered long ARG (sampled to mid-2022) that represent previously unnamed putative recombinant sequences, which do not have a Pango recombinant designation. For this proof-of-concept study, we have arbitrarily picked 12 recombination nodes that seem plausible (Supplementary Table 4). Coincidentally, all these nodes involve Omicron subvariants (BA.1, BA.2, BA.4, and BA.5). These recombination nodes (1) were inserted into the long ARG on or after January 8, 2022; (2) have at least 10 descendent samples; and (3) have no mutational differences (including immediate reversions) from their parent nodes.

TODO: Node  740761 (USA/NC-CDC-LC0668306/2022) is a sister clade proposed1006 in the UShER public phylogeny.
Node 628656 (Scotland/QEUH-37794BB/2022) is a sister clade of XAC, and annotated as “miscBA2BA1PostSpike”.

\subsection{Recombinant origins of the Pango lineages XA and XAG}

XX

\subsection{Distribution of recombination breakpoint intervals}

We explore recombination signals across the SARS-CoV-2 genome by inspecting the breakpoint intervals of the filtered recombination nodes from the wide ARG (sampled to mid-2021) (Figure 4). The length of a breakpoint interval reflects uncertainty in the estimated location of a breakpoint. On average, the length of the breakpoint intervals is 2,895 bases (median, 1,670 bases; range, 1 to 23,666 bases). This broad distribution of interval length indicates that the location of a breakpoint often could not be precisely determined given the observed genetic variation.

There is a preponderance of breakpoint intervals spanning the 3’ end of the genome (Figure 4A), as noted previously by Turakhia et al. \cite{Turakhia2022-it}. Inspecting more closely, we find an enrichment in the number of intersecting intervals in the E gene, ORF8, ORF8-N intergenic region, and 3’ untranslated region (Figure 4B). Markedly, there is an abrupt drop in the number of intersecting intervals in the N gene following the peak in ORF8 towards the 3’ end.

Next, we investigate the potential recombination signals alongside the number of mutations across the genome (Figure 4C). Regions with a higher density of mutations contain more genetic variation, increasing the power to detect recombination events. The position 28,271 (located in the ORF8-N intergenic region) has a particularly high number of mutations in the samples (Figure 5). At this position, mutations (n = 7,572), insertions (n = 5,782), and deletions (n = 1,605) appear to occur frequently. One possible explanation is that this homopolymeric region is prone to sequencing errors, giving rise to an artifactual recombination signal. At the upstream position 27,972, numerous mutations also occur. This position has been reported to be hypermutable \cite{Jungreis2021-dh}.

We further consider the breakpoint intervals within the context of recombination breakpoint sequence motifs. It is hypothesised that certain palindromic breakpoint sequences (specifically, CAGAC and CAGAT) promote template switching during replication in SARS-CoV-2 via formation of base-paired stem loops in the genome structure \cite{Gallaher2020-lb}. There are 87 occurrences of these two breakpoint sequences in the Wuhan-Hu-1/2019 reference sequence (Supplementary Table 5). We observe that XX (XX\%) of the breakpoint intervals of the HMM-consistent recombinants span at least one of the breakpoint sequences.

\subsection{Mutational spectrum}

In addition to recombination events, the ARGs record mutation events along the SARS-CoV-2 phylogeny. Yi et al. \cite{Yi2021-sc} reconstructed a SARS-CoV-2 phylogeny of over 350,000 genomes sampled globally from December 24, 2019, to January 12, 2021, and systematically catalogued all the mutations occurring along the phylogeny. Notably, in the mutational spectrum described by Yi et al., C-to-U and G-to-U mutations occur more frequently than U-to-C and U-to-G mutations, respectively.

For validation purposes, we compare the mutational spectrum from the wide ARG with the mutational spectrum of Yi et al. We categorised all of the single nucleotide mutations by type (defined by the ancestral state and derived state), excluded mutations inherited by only a single sample (which are enriched for sequencing errors and artefacts), and tallied them up by type. Similarly, we took the data for single nucleotide mutations from Yi et al. \ref{https://github.com/ju-lab/SC2_evol_signature}, excluded mutations occurring along terminal branches, and tallied them up by type. We find that the mutational spectrum from the wide ARG (448,825 mutations) matches the mutational spectrum reported by Yi et al. (92,344 mutations) (Figure X). In both spectra, C-to-U mutations and G-to-U mutations occur more frequently than U-to-C and U-to-G, respectively (Figure X). Similar results are obtained when including the mutations inherited by only a single sample or occurring on terminal branches (data not shown).

\subsection{Mutations under selection}

XX

\section{Discussion}

The COVID-19 pandemic continues to be a global emergency, with persistent high levels of infection. This high prevalence has allowed the proliferation of many variants, with more than 600 Pango-designated lineages circulating globally in the last three months (January to March, 2023; https://gisaid.org/; access on March 27, 2023). High prevalence also brings a higher risk of coinfection, increasing opportunities for new phenotypically distinct recombinants to emerge and spread.

In this proof-of-concept study, we demonstrate the feasibility of reconstructing a large ARG from over one million SARS-CoV-2 genomes, a feat that is simply not possible using other available ARG inference methods (Rasmussen et al. 2014; Ignatieva et al. 2021; Speidel et al. 2019). We have accomplished this feat by building on top of recent advances in storing and processing large ARGs (tskit) and inferring genealogical relationships efficiently at scale (tsinfer). The highly compact representation of the ARGs as succinct tree sequences, coupled with vectorised computing technologies, enables the large SARS-CoV-2 ARGs to be loaded in less than one second in a Jupyter notebook and allows for each SARS-CoV-2 genome to be stored in less than eleven bytes of data. We believe that further development of the sc2ts methodology and software can create a powerful tool (1) to efficiently detect emerging or previously overlooked recombinants in SARS-CoV-2; (2) to jointly model the mutational and recombinational processes shaping global SARS-CoV-2 genomic diversity; and (3) to perform a broad variety of evolutionary analyses of SARS-CoV-2 genomes with unprecedented global sequencing coverage.

ARGs are rich genealogical structures that contain information about the genetic relationships among sampled genomes in the presence of recombination. Here, we have conducted several analyses of the preliminary ARGs to show that the ARGs built using sc2ts can (1) recapitulate the phylogenetic relationships among major SARS-CoV-2 lineages; (2) recover biologically plausible recombination signals in SARS-CoV-2 genomes; and (3) model mutation events in SARS-CoV-2 evolutionary history.

ARGs reconstructed using sc2ts capture the phylogenetic relationships of major SARS-CoV-2 lineages. Our results show that the backbone topology of a densely sampled wide ARG inferred using sc2ts agrees well with the backbone topology of a NextStrain tree inferred using IQ-Tree2, a widely used phylogenetic inference package. However, we have noted several areas of differences (e.g., poorly calibrated branch lengths), which are the focus of future development of sc2ts. We believe that some of these differences can be reduced by improving and refining our approach and strategy to handle samples with erroneously reported collection dates (i.e., time travellers that have passed the maximum submission delay of 30 days) and to modify the topology of daily trees.

Our results show that sc2ts identifies the early recombinants (pre-2022) previously identified by Jackson et al. (2021). Sc2ts also identifies many Pango recombinants, which are designated after community perusal of supporting genomic evidence. We have found several previously unreported and overlooked recombinant sequences in the wide ARG. After further improvements and enhancements, we will use sc2ts to build a comprehensive ARG to search for recombination signals among all the 15 million and counting SARS-CoV-2 genomes available.

“The rate of putative recombination breakpoints is about three times higher towards the 3' of the change point than the 5' interval.“ Consistent with this, the breakpoint intervals of the putative recombinants identified here tend to occur on the 3’ end of the SARS-CoV-2 genome.

\section{Improvements}

Currently, sc2ts is in development. Below, we propose several ways to potentially improve or enhance sc2ts. We welcome suggestions and ideas from experts.

\begin{itemize}
\item A potential source of error in sc2ts ARGs comes from incorrectly reported collection dates of the samples, which can cause sample genomes to be attached to the wrong parts of the ARGs. Thus far, we have attempted to correct the most offending errors in the ARGs by filtering out "time travellers", but the sampling time can still be wrong by days or tens of days. ***describe possible approaches to correct this?***
\item To avoid time travellers, we used a maximum submission delay, but this will systematically bias against submissions from countries with greater submission delays (Kalia et al.). An area of future work is to add back the excluded sequences to a built ARG to assess which ones are likely true time travellers.
\item To avoid time travellers, we used a maximum submission delay, but this will systematically bias against submissions from countries with greater submission delays (Kalia et al.). An area of future work is to add back the excluded sequences to a built ARG to assess which ones are likely true time travellers.
\item We have used a single mismatch value in all the HMM runs, set so that having three mutations is as likely as having a single recombination. This parameter choice is arbitrary, and higher values may reduce the number of origins for those Pango lineages that were inferred to descend from more than one recombinant ancestor. In addition, because both the mutation rate and recombination rate varies across the SARS-CoV-2 genome, it may be helpful to allow the mismatch value to vary along the genome.
\item By treating mutations equally in the current HMM implementation, we implicitly assume that the rates of different types of mutations are identical. However, as reported previously (Yi et al. 2021), there is a high degree of mutational asymmetry in SARS-CoV-2. To account for mutational biases, more sophisticated nucleotide substitution models (e.g., allowing for different rates of transition and transversion) can be incorporated into HMM likelihood calculations.
\item Our current masking strategy handles indels as missing data. Doing this can remove information helpful to detect recombinants. For example, sequences of the Pango recombinant XP were not identified as recombinants, because XP’s characteristic multibase deletion was masked.
\item This "lack of memory" in the HMM can create recombinants with three or more parents when it would be equally parsimonious to have just two. The copying process in the HMM currently does not limit the number of parents to two, but doing so would be a valuable addition.
\item Some recombinants may reasonably have three or more parents, indicating multiple "stacked" recombination events. We have largely ignored these in the current analysis, but with further sc2ts development, there is potential to analyse these putative recombinants and trace their evolutionary origins through the ARG.
\item In the present implementation, HMM traceback is performed once, yielding a single maximum-likelihood genome copying path. A stochastic traceback algorithm (e.g., (Rasmussen et al. 2014)) can be implemented in order to quantify uncertainty about putative parent node(s) and switch positions. These estimates of uncertainty may be used as support values for branches analogous to bootstrap values produced by standard phylogenetic methods.
\item Daily trees are built using the UPGMA algorithm. This entails an extra step to “un-resolve” daily trees in parts of the trees with no phylogenetic resolution. To avoid this step, a better approach is to use a maximum-parsimony algorithm to build daily trees instead.
\end{itemize}

\section{Conclusions}

The global understanding of viral evolution during the COVID-19 pandemic is unprecedented, with over 15 million sequences shared to date. We envisage that once fully developed and tested, sc2ts methodology and software may provide a unified computational framework to study the evolution and epidemiology of SARS-CoV-2 at pandemic scale. The evolutionary analyses presented herein exemplify the numerous applications possible using the ancestral recombination graph inferred by sc2ts. We invite bioinformaticians, evolutionary biologists, and virologists to join us and the broader tskit community to build the most complete representation of the evolutionary history of SARS-CoV-2. Moreover, while we have focussed on the particular application to SARS-CoV-2, the methods we describe are readily applicable to the analysis of many other viral pathogens. Solving the challenges in building and analysing large viral genealogies would therefore contribute to the important goal of improving preparedness for potential future pandemics.

\section{Acknowledgements}

We gratefully acknowledge all data contributors, i.e., the Authors and their Originating laboratories responsible for obtaining the specimens, and their Submitting laboratories for generating the genetic sequence and metadata and sharing via the GISAID Initiative, on which this research is based. Also, we thank Dr. Morag Graham (the National Microbiology Laboratory, Winnipeg, Canada) for kindly providing the genomic coordinates of the breakpoint sequence motifs. SHZ is supported by the Janssen-Oxford Translational Genomics Fellowship. JK, YW, and BJ are supported by the Robertson Foundation.

\begin{table}
\centering
\begin{tabular}{l|l|l}
\hline
\textbf{Property} & \textbf{Wide ARG} & \textbf{Long ARG} \\
\hline
Last sampling date & 2021-06-30 & 2022-06-30 \\
Local trees & 1,496 & 958 \\
\hline
\multicolumn{3}{l}{\textit{Nodes}} \\
All nodes (sample and non-sample) & 1,453,347 & 783,231 \\
Sample nodes & 1,265,685 (87\%) & 657,239 (84\%) \\
Tree adjustment nodes (excluding recombination nodes) & 80,830 (6.34\%) & 72,107 (10.97\%) \\
Recombination nodes & 4,123 (0.28\%) & 2,078 (0.27\%) \\
\hline
\multicolumn{3}{l}{\textit{Types of recombination nodes}} \\
Singletons (i.e. no descendant besides the sequence causing insertion of the node) & 2,601 (63\%) & 1,315 (63\%) \\
Two-parent recombinants & 3,521 (85\%) & 1,869 (90\%) \\
HMM-consistent recombinants & XX (XX\%) & XX (XX\%) \\
\hline
\multicolumn{3}{l}{\textit{Mutations}} \\
Total mutations & 1,213,193 & 1,062,072 \\
Mutations on a terminal edge & 758,903 (62.55\%) & 767,111 (72.23\%) \\
Recurrent mutations & 74,719 (6.16\%) & 50,099 (4.72\%) \\
Reversions & 72,617 (5.99\%) & 48,226 (4.54\%) \\
Immediate reversions (one intervening node only) & 3,541 (0.29\%) & 1,347 (0.13\%) \\
\hline
\multicolumn{3}{l}{\textit{Storage and loading time}} \\
tszipped file (including sample metadata, imputed Pango lineage assignments, and results of HMM runs for recombination nodes) & ~57 Mb & ~36 Mb \\
Bytes to encode one sample & 8.29 & 10.83 \\
Load time (seconds) & 0.9 & 0.5 \\
\hline
\end{tabular}
\end{table}

\begin{table}
\centering
\begin{tabular}{l|c|c|c|c|c}
\hline
\multicolumn{1}{c}{} & \multicolumn{3}{c}{\textbf{Jackson et al. (2021)}} & \multicolumn{2}{c}{\textbf{sc2ts (wide ARG)}} \\
\hline
\textbf{Sample} & \textbf{Group} & \textbf{Parents} & \thead{Breakpoint \\ interval(s)} & \textbf{Parents} & \thead{Breakpoint \\ interval(s)} \\
\hline
ALDP-11CF93B & A & \thead{B.1.177 \\ Alpha} & 21,255–21,613 & \thead{B.1.177.18 \\ Alpha} & 21,256–22,227 \\
ALDP-125C4D7 & A & \thead{B.1.177 \\ Alpha} & 21,255–21,613 & \thead{B.1.177.18 \\ Alpha} & 21,256–22,227 \\
ALDP-130BB95 & A & \thead{B.1.177 \\ Alpha} & 21,255–21,613 & \thead{B.1.177.18 \\ Alpha} & 21,256–22,227 \\
LIVE-DFCFFE & A & \thead{B.1.177 \\ Alpha} & 18,998–20,294 & \thead{B.1.177.18 \\ Alpha} & 21,256–22,227 \\
QEUH-CCCB30 & B & \thead{B.1.36.28 \\ Alpha} & 6,528–6,953 & \thead{B.1.36 \\ Alpha} & 6,529–6,954 \\
QEUH-CD0F1F & B & \thead{B.1.36.28 \\ Alpha} & 6,528–6,953 & \thead{B.1.36 \\ Alpha} & 6,529–6,954 \\
MILK-1166F52 & C & \thead{Alpha \\ B.1.221.1} & 25,996–27,441 & \thead{Alpha \\ B.1.221} & 25,997–27,972 \\
MILK-11C95A6 & C & \thead{Alpha \\ B.1.221.1} & 25,996–27,441 & \thead{Alpha \\ B.1.221} & 25,997–27,972 \\
QEUH-109B25C & C & \thead{Alpha \\ B.1.221.1} & 25,996–27,441 & \thead{Alpha \\ B.1.221} & 25,997–27,972 \\
MILK-126FE1F & D & \thead{B.1.36.39 \\ Alpha} & 20,703–23,062 & \thead{B.1.36.39 \\ Alpha} & 22,445–23,063 \\
RAND-12671E1 & D & \thead{B.1.36.39 \\ Alpha} & 20,703–23,062 & \thead{B.1.36.39 \\ Alpha} & 22,445–23,063 \\
RAND-128FA33 & D & \thead{B.1.36.39 \\ Alpha} & 20,703–23,062 & \thead{B.1.36.39 \\ Alpha} & 22,445–23,063 \\
CAMC-CBA018 & n/a & \thead{B.1.177 \\ Alpha} & 20,389–21,254 & \thead{B.1.177 \\ Alpha} & 16,177–21,255 \\
CAMC-CB7AB3 & n/a & \thead{Alpha \\ B.1.177 \\ Alpha} & \thead{3,267–4,474 \\ 20,389–21,254} & \thead{Alpha \\ B.1.177 \\ Alpha} & \thead{3,268–5,388 \\ 16,177–21,255} \\
MILK-103C712 & n/a & \thead{B.1.177.17 \\ Alpha} & \thead{408–444 \\ 26,801–27,876} & n/a & n/a \\
QEUH-1067DEF & n/a & \thead{Alpha \\ B.1.177.9} & 10,523–10,869 & \thead{Alpha \\ B.1.177} & 7,729–10,870 \\
\hline
\end{tabular}
\caption{\label{tab:jackson}Recombinant sequences involving the Alpha (B.1.1.7) variant reported by Jackson et al. \cite{Jackson2021-ik} have recombinant ancestry in the wide ARG. The breakpoint intervals and Pango lineage assignments of the parents were taken from Table 2 (3SEQ results) of Jackson et al., except the Pango lineage assignment of the parents of group B recombinants, which were taken from Table 1 (motif-based results).}
\end{table}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{figures/overview_sc2ts.pdf}
\caption{\label{fig:overview_sc2ts} A schematic of the sc2ts method. Sc2ts reconstructs the genetic relationships among SARS-CoV-2 genomes by copying samples to all possible ancestors collected at earlier time points (curved arrows). An ARG is built by attaching new sample genomes grouped by collection date to an existing ARG rooted at the Wuhan-Hu-1 reference genome, day by day. Each iteration involves three stages: (1) attachment of new samples to the growing ARG (A, B); (2) reconstruction of trees relating the samples under each attachment node (C); and parsimony-based tree topology adjustments (D, E). In the absence of  recombination, sc2ts infers an ARG that is a single tree relating the samples (A). However, in the presence of recombination, sc2ts infers an ARG that is encoded as a sequence of local trees relating segments of the sample genomes (B), inserting a recombination node to represent a putative recombination event. Additionally, mutation-collapsing nodes (D) and pushed-reversion nodes (E) are inserted to make more parsimonious placements of mutations that should be shared or should not be immediately reverted, respectively.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/ls_diagram_covid.pdf}
\caption{\label{fig:ls_diagram} A schematic of the Li and Stephens (LS) model. The LS model in its implementation as an HMM is governed by the transition and emission matrices, \textit{Q} and \textit{E}, respectively. Each may be a function of the reference panel members, but transitions are generally assumed to be independent of the hidden states themselves (shown in the pink panel). This dramatically increases performance as the state space drops to two states (i.e., switching or not switching). The recombination probability at site \textit{l} is \textit{rl}, while n denotes the number of members in the reference panel. Emissions may also be a function of the nucleotide states, but in our case, we assume that all observed alleles (in both the reference panel and focal sequence, \textit{al}) at the site under consideration are equally likely and occur with mutation probability \textit{l} (shown in the blue panel). These two processes, transition and emission (which encode recombination and mutation, respectively), define the generative process of the HMM, by which an imperfect mosaic is emitted from a reference panel of samples and nodes from the ARG inserted at earlier collection dates (shown in the lower panel). Here, the unseen states in the reference panel are shown as coloured lines enclosed by the grey box, and the true path through the data which leads to the emitted focal sequence is displayed by the black arrow. Examples of transition and emission probabilities along this trajectory are shown by the red and blue arrows, respectively. Black crosses along the focal sequence show genotyping errors or mutations. Note that in practice, we do not know the pattern of colours along the focal sequence and must infer it using the Viterbi algorithm.}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth]{figures/mutational_spectra.pdf}
\caption{\label{fig:mutational_spectra} Mutational spectrum of SARS-CoV-2 in the wide ARG (sampled to mid-2021) compared to that reported by Yi et al. \cite{Yi2021-sc}. Mutations are categorised by type (i.e., ancestral state > derived state). The percentages of each mutation type from the wide ARG are represented by blue bars and the percentages from Yi et al. by orange bars, with the darker colours representing one direction (e.g., C > U) and the lighter colours the reverse (e.g., U > C).}
\end{figure}

\bibliographystyle{alpha}
\bibliography{references}

\end{document}